### 2020-09-18 presentations/bsj_2020-09-19/index.Rmd
Update BSJ presentation

```{r, load-refs, includ = FALSE, cache = FALSE}
# Set bibliography options
library(RefManageR)
BibOptions(
  ignore.case = TRUE,
  check.entries = FALSE,
  max.names = 3,
  bib.style = "authoryear", # set style of bibliography as e.g., "Nitta 2020..."
  style = "text" # set style of citations as plain text (no links)
)

# Read in bibliography
bib <- ReadBib(here::here("presentations/bsa_2020-07-27/refs.bib"), check = FALSE)
```

@@

# Load targets
loadd(list = c(
  "biodiv_ferns_spatial",
  "occ_point_data_ferns",
  "protected_areas",
  "traits_for_dist",
  "comm_ferns",
  "green_list",
  "ppgi",
  "japan_shp"),
  cache = ja_fern_cache)

@@

# Save currently used data as back-up in case data produced by plan changes
data <- list(
  biodiv_ferns_spatial = biodiv_ferns_spatial,
  occ_point_data_ferns = occ_point_data_ferns,
  protected_areas = protected_areas,
  traits_for_dist = traits_for_dist,
  comm_ferns = comm_ferns,
  green_list = green_list,
  ppgi = ppgi,
  japan_shp = japan_shp,
  collections_by_country = collections_by_country
)

saveRDS(data, here::here("presentations/bsj_2020-09-19/data/data.RDS"))

# Change cluster to factor
biodiv_ferns_spatial <- biodiv_ferns_spatial %>%
  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
  # Add redundancy
  mutate(redundancy = 1 - (richness/abundance))

@@

# Set CRS
japan_crs <- sf::st_crs(japan)

biodiv_ferns_spatial <- sf::st_set_crs(biodiv_ferns_spatial, japan_crs)
```

@@

.pull-left-wide[
# 日本のシダ植物多様性の<br>多次元解析
<br>
## 新田ジョエル<sup>1</sup>、海老原淳<sup>2</sup>

<sup>1</sup>東京大学理学系研究科・岩崎研究室<br>
<span style = 'font-size: 80%;'>https://joelnitta.com</span>

<sup>2</sup>国立科学博物館・植物研究部

日本植物学会第84回大会<br><span style = 'font-size: 80%;'>2020.03.01</span>
]

@@

a + b + plot_layout(widths = c(3,1), heights = c(1,1))
```

@@

```{r pe, out.width = "120%", fig.asp = 0.7}
ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
  scale_fill_manual(values = canape_cols) +
  labs(fill = "Endemism type") +
  map_theme() +
  theme(
    legend.title = element_markdown(),
    legend.position = "right"
  )
```

@@

---
## ほとんどの地点は四つのバイオリージョンのどれかに所属する

@@

---
## 系統的距離を使うと琉球と小笠原が一つのバイオリージョンになる

@@

>>>>

```{r, load-refs, includ = FALSE, cache = FALSE}
# Set bibliography options
library(RefManageR)
BibOptions(
  ignore.case = TRUE,
  check.entries = FALSE,
  max.names = 3,
  bib.style = "authoryear", # set style of bibliography as e.g., "Nitta 2020..."
  style = "text" # set style of citations as plain text (no links)
)
```

@@

# Load targets
loadd(list = c(
  "biodiv_ferns_spatial",
  "occ_point_data_ferns",
  "protected_areas",
  "traits_for_dist",
  "comm_ferns",
  "green_list",
  "ppgi",
  "japan_shp",
  "signif_cells_protected_area"),
  cache = ja_fern_cache)

@@

# Read in bibliography
bib <- ReadBib(here::here("presentations/bsa_2020-07-27/refs.bib"), check = FALSE)

# Read in Green List in Japanese
green_list_ja <- readxl::read_excel(here::here("presentations/bsj_2020-09-19/data/FernGreenListV1.01.xls")) %>%
  transmute(taxon_id = as.character(ID20160331), name_ja = 新リスト和名)

# Save currently used data as back-up in case data produced by plan changes
data <- list(
  biodiv_ferns_spatial = biodiv_ferns_spatial,
  occ_point_data_ferns = occ_point_data_ferns,
  protected_areas = protected_areas,
  traits_for_dist = traits_for_dist,
  comm_ferns = comm_ferns,
  green_list = green_list,
  ppgi = ppgi,
  japan_shp = japan_shp,
  collections_by_country = collections_by_country,
  bib = bib,
  green_list_ja = green_list_ja,
  signif_cells_protected_area = signif_cells_protected_area
)

saveRDS(data, here::here("presentations/bsj_2020-09-19/data/data.RDS"))
```

```{r prep-data}
# Change cluster to factor
biodiv_ferns_spatial <- biodiv_ferns_spatial %>%
  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
  # Add redundancy
  mutate(redundancy = 1 - (richness/abundance))

@@

# Set CRS
st_crs(japan) <- 4326
st_crs(biodiv_ferns_spatial) <- 4326
st_crs(protected_areas) <- 4326
```

@@

.pull-left-wide[
# 日本のシダ植物多様性の<br>多次元解析
<br>
## 新田ジョエル<sup>1</sup>・Brent Mishler<sup>2</sup><br>岩崎渉<sup>1</sup>・海老原淳<sup>3</sup>

<sup>1</sup>東京大学理学系研究科<br>

<sup>2</sup>University of California, Berkeley

<sup>3</sup>国立科学博物館

日本植物学会第84回大会<br><span style = 'font-size: 80%;'>2020.03.01</span>

<span style = 'font-size: 80%;'>https://joelnitta.com</span>

]

@@

a + b + plot_layout(widths = c(3,1), heights = c(1,1))

ggsave(here::here("presentations/bsj_2020-09-19/richness.png"), height = 6, width = 10, units = "in")
```

@@

```{r pe, out.width = "120%", fig.asp = 0.7}
biodiv_ferns_spatial_simple_endem <-
  biodiv_ferns_spatial %>%
  mutate(endem_type = case_when(
    endem_type == "super" ~ "mixed",
    TRUE ~ as.character(endem_type)
  )) %>%
  mutate(endem_type = factor(endem_type, levels = c("neo", "paleo", "mixed", "not significant")))

# CANAPE
canape_cols_simple <-
  c(
    "neo" = "#e31a1c",
    "paleo" = "#1f78b4",
    "mixed" = "#6a51a3",
    "not significant" = "#ffffcc"
  )

ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial_simple_endem, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
  scale_fill_manual(values = canape_cols) +
  labs(fill = "Endemism type") +
  map_theme() +
  theme(
    legend.title = element_markdown(),
    legend.position = "right"
  )

ggsave(here::here("presentations/bsj_2020-09-19/canape_simple.png"), height = 6, width = 6, units = "in")

```

@@

---
## 主なバイオリージョンが四つある

@@

---
系統的距離を使うと琉球と小笠原が一つのバイオリージョンになる

@@

---
background-position: 50% 50%
background-image: url(images/species-background-1.png)
background-size: contain

## 主なバイオリージョンが四つある

```{r find-top-species}
# Extract clusters
clusters <-
  biodiv_ferns_spatial %>%
  as_tibble() %>%
  select(site = "grids", cluster = phylo_cluster, -geometry)

# Convert community df to long format
comm_ferns_long <-
  comm_ferns %>%
  rownames_to_column("site") %>%
  pivot_longer(names_to = "taxon", values_to = "abun", -site)

# Find top-scoring species (with high occurrence inside one cluster, and low elsewhere)
top_species <-
  comm_ferns_long %>%
  left_join(clusters, by = "site") %>%
  filter(cluster %in% 1:4) %>%
  group_by(taxon, cluster) %>%
  summarize(
    abun = sum(abun),
    total = n()
  ) %>%
  mutate(
    # percent abundance within a bioregion
    perc_abun = abun / total
  ) %>%
  arrange(taxon, desc(perc_abun)) %>%
  select(taxon, cluster, perc_abun) %>%
  group_by(taxon) %>%
  # take top two highest percent abundances per species, calculate their difference
  slice_max(n = 2, order_by = perc_abun, .preserve = TRUE, with_ties = FALSE) %>%
  mutate(
    top_cluster = cluster[perc_abun == max(perc_abun)]
  ) %>%
  summarize(
    max_dif = max(perc_abun) - min(perc_abun),
    cluster = unique(top_cluster)
  ) %>%
  # keep only those with greater than 65% difference between top 1st and 2nd
  # most abundant bioregions
  filter(max_dif > 0.5) %>%
  group_by(cluster) %>%
  # Choose top three per cluster for plotting
  slice_max(n = 3, order_by = max_dif, with_ties = FALSE) %>%
  # Add name in Japanese
  left_join(select(green_list, taxon, taxon_id), by = "taxon") %>%
  left_join(green_list_ja, by = "taxon_id")
```

```{r plot-top-species-prep}
# make helper function to plot top-scoring species
plot_top_taxon <- function (taxon_select, taxon_print, biodiv_ferns_spatial, comm_ferns_long) {
biodiv_ferns_spatial %>%
  left_join(
    filter(comm_ferns_long, taxon == taxon_select),
    by = c(grids = "site")
  ) %>%
  mutate(abun = factor(abun)) %>%
  ggplot(aes(fill = abun)) +
  scale_fill_manual(values = c("#404040", "#d01c8b")) +
  geom_sf(color = "transparent") +
  labs(subtitle = taxon_print) +
  simple_map_theme_light() +
  theme(plot.subtitle = element_text(family = "HiraKakuPro-W3"))
}
```

```{r plot-top-species-cluster-1, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial %>%
  mutate(focus = ifelse(taxonomic_cluster == 1, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  simple_map_theme_light() +
  labs(subtitle = "Region 1")

focus_region +
plot_top_taxon(top_species$taxon[[1]], taxon_print = top_species$name_ja[[1]], biodiv_ferns_spatial, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[2]], taxon_print = top_species$name_ja[[2]], biodiv_ferns_spatial, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[3]], taxon_print = top_species$name_ja[[3]], biodiv_ferns_spatial, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave(here::here("presentations/bsj_2020-09-19/plot-top-species-cluster-1.png"), height = 6, width = 6, units = "in")

# One species per region
plot_top_taxon(top_species$taxon[[2]], taxon_print = top_species$name_ja[[2]], biodiv_ferns_spatial, comm_ferns_long) + 
plot_top_taxon(top_species$taxon[[4]], taxon_print = top_species$name_ja[[4]], biodiv_ferns_spatial, comm_ferns_long) + 
plot_top_taxon(top_species$taxon[[7]], taxon_print = top_species$name_ja[[7]], biodiv_ferns_spatial, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave(here::here("presentations/bsj_2020-09-19/plot-top-species-one-each-phylo.png"), height = 6, width = 6, units = "in")
```

.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]

---
background-position: 50% 50%
background-image: url(images/species-background-2.png)
background-size: contain

## 主なバイオリージョンが四つある

```{r plot-top-species-cluster-2, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial %>%
  mutate(focus = ifelse(taxonomic_cluster == 2, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  simple_map_theme_light() +
  labs(subtitle = "Region 2")

focus_region +
plot_top_taxon(top_species$taxon[[4]], taxon_print = top_species$name_ja[[4]], biodiv_ferns_spatial, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[5]], taxon_print = top_species$name_ja[[5]], biodiv_ferns_spatial, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[6]], taxon_print = top_species$name_ja[[6]], biodiv_ferns_spatial, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave(here::here("presentations/bsj_2020-09-19/plot-top-species-cluster-2.png"), height = 6, width = 6, units = "in")
```

.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]

---
background-position: 50% 50%
background-image: url(images/species-background-3.png)
background-size: contain

## 主なバイオリージョンが四つある

```{r plot-top-species-cluster-3, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial %>%
  mutate(focus = ifelse(taxonomic_cluster == 3, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  simple_map_theme_light() +
  labs(subtitle = "Region 3")

focus_region +
plot_top_taxon(top_species$taxon[[7]], taxon_print = top_species$name_ja[[7]], biodiv_ferns_spatial, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[8]], taxon_print = top_species$name_ja[[8]], biodiv_ferns_spatial, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[9]], taxon_print = top_species$name_ja[[9]], biodiv_ferns_spatial, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave(here::here("presentations/bsj_2020-09-19/plot-top-species-cluster-3.png"), height = 6, width = 6, units = "in")
```

.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]

---
background-position: 50% 50%
background-image: url(images/species-background-4.png)
background-size: contain

## 主なバイオリージョンが四つある

```{r plot-top-species-cluster-4, out.width = "120%", fig.asp = 0.65}
bonin_box <-
biodiv_ferns_spatial %>%
  filter(taxonomic_cluster == 4) %>%
  sf::st_bbox()

expand_value <- 0.4

focus_region <-
biodiv_ferns_spatial %>%
  mutate(focus = ifelse(taxonomic_cluster == 4, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  simple_map_theme_light() +
  labs(subtitle = "Region 4") +
  annotate(
    "rect",
           xmin = bonin_box[["xmin"]] - expand_value,
           xmax = bonin_box[["xmax"]] + expand_value,
           ymin = bonin_box[["ymin"]] - expand_value,
           ymax = bonin_box[["ymax"]] + expand_value,
    fill = "transparent",
    color = "grey10")

focus_region +
plot_top_taxon(top_species$taxon[[10]], taxon_print = top_species$name_ja[[10]], biodiv_ferns_spatial, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[12]], taxon_print = top_species$name_ja[[12]], biodiv_ferns_spatial, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[11]], taxon_print = top_species$name_ja[[11]], biodiv_ferns_spatial, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave(here::here("presentations/bsj_2020-09-19/plot-top-species-cluster-4.png"), height = 6, width = 6, units = "in")
```

.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]

---
## バイオリージョンは従来の植生帯と概ね一致

.pull-left[
Taxonomic bioregions

```{r tax-bioregion-forest}
ggplot(biodiv_ferns_spatial, aes(fill = taxonomic_cluster)) +
  geom_sf(color = "transparent") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp_light() +
  theme(legend.position = "none")
```
]

.pull-right[
植生帯<br>
.small[`r Citet(bib, "Shimizu2014")`<br><br>]
<img src="images/forest_types_eng.png" alt="drawing" width="280"/>
]

# 生物多様性はどれくらい守られているのか？

# How well is biodiversity protected?

---
## 日本は一般的によく守られている（平均30%）

```{r protected-areas-graph}
# Calculate the total area of Japan
area_of_japan <- 377900

# Calculate baseline protection percentage: percent protected areas across the whole country
total_percent_protected <-
  protected_areas %>%
  as_tibble() %>%
  select(-geometry) %>%
  filter(status %in% c("medium", "high")) %>%
  group_by(status) %>%
  summarize(
    total_area = sum(area),
    .groups = "drop"
  ) %>%
  mutate(percent_protected = total_area/area_of_japan) %>%
  mutate(percent_protected = as.numeric(percent_protected))

# Extract values for medium and high protection level
total_percent_protected_medium <- total_percent_protected %>% filter(status == "medium") %>% pull(percent_protected)
total_percent_protected_high <- total_percent_protected %>% filter(status == "high") %>% pull(percent_protected)
total_percent_protected_all <- sum(total_percent_protected_medium, total_percent_protected_high)

show_col(viridis_pal(option = "D")(3))

protection_cols <- c(
  Medium = "#21908CFF",
  High = "#FDE725FF"
)

# Plot protected areas
signif_cells_protected_area %>%
  mutate(
    metric = case_when(
      metric == "pd" ~ "PD",
      metric == "fd" ~ "FD",
      metric == "pe" ~ "PE",
      metric == "richness" ~ "Richness",
    ),
    status = str_to_sentence(status)) %>%
  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "Richness"))) %>%
  ggplot(aes(x = percent_protected, y = metric, fill = status)) +
  geom_col(position = position_dodge()) +
  geom_vline(xintercept = total_percent_protected_medium, color = "grey40", linetype = 3) +
  geom_vline(xintercept = total_percent_protected_high, color = "grey40", linetype = 2) +
  scale_fill_manual(values = protection_cols) +
  scale_x_continuous(labels = scales::percent) +
  labs(
    x = "Percent protected",
    y = "Biodiversity metric",
    fill = "Conservation\nstatus"
  ) +
  theme_bw(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "right"
  )

ggsave(here::here("presentations/bsj_2020-09-19/protected_bar.png"), height = 6, width = 8, units = "in")
```
